European or American? A re-examination of the transatlantic influence over the IMF

### 2025-09-11 manuscript/manuscript.qmd
fix 10th para
remove the para on 3 european powers

Second, our measures of influence, while grounded in the existing literature, capture alignment rather than direct lobbying or pressure. We cannot observe the specific mechanisms through which influence is exercised or the content of behind-the-scenes negotiations. Qualitative case studies examining specific lending decisions could provide valuable complementary insights into the micro-foundations of the patterns we document.

Third, our focus on the three largest European powers—the UK, France, and Germany—may understate the collective influence of the European Union as a whole. As EU coordination mechanisms develop and potentially consolidate European representation within international financial institutions, the dynamics we analyse may continue to evolve. Research examining the role of EU institutional actors, such as the European Central Bank or European Commission, in IMF decision-making would provide a useful extension of our work.

Fourth, our analysis focuses exclusively on lending outcomes and does not examine other dimensions of IMF operations, such as surveillance, technical assistance, or institutional governance. European and American influence may manifest differently across these various functions. Similarly, our study does not address the effectiveness or appropriateness of the lending arrangements influenced by great power politics—a question that would require additional data on economic outcomes and counterfactual analysis.

>>>>

Second, our measures of influence, while grounded in the existing literature, capture alignment rather than direct lobbying or pressure. We cannot observe the specific mechanisms through which influence is exercised or the content of behind-the-scenes negotiations. Qualitative case studies examining specific lending decisions could provide valuable complementary insights into the micro-foundations of the patterns we document. However, such approaches often rely on extensive insider information and interviews with IMF officials, which are difficult to obtain given the Fund’s highly consensus-based and closed-door decision-making culture.

Third, our analysis focuses exclusively on lending outcomes and does not examine other dimensions of IMF operations, such as surveillance, technical assistance, or institutional governance. European and American influence may manifest differently across these various functions. Similarly, our study does not address the effectiveness or appropriateness of the lending arrangements influenced by great power politics—a question that would require additional data on economic outcomes and counterfactual analysis.
